Non-word repetition in Yélî Dnye

### 2020-09-11 manuscript.tex

\documentclass[english,,man,floatsintext]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\PassOptionsToPackage{hyphens}{url} % url is loaded by hyperref
\usepackage[unicode=true]{hyperref}
\hypersetup{
            pdftitle={Non-word repetition in Yélî Dnye},
            pdfauthor={Alejandrina Cristia~\& Marisa Casillas},
            pdfkeywords={phonology, non-word repetition, development},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[shorthands=off,main=english]{babel}
\else
  \usepackage{polyglossia}
  \setmainlanguage[]{english}
\fi
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

\setmainfont{Doulos SIL} % Set main font to Doulos SIL

\title{Non-word repetition in Yélî Dnye}
\author{Alejandrina Cristia\textsuperscript{1}~\& Marisa
Casillas\textsuperscript{2}}
\date{}

\authornote{All data are made available in a repository in the
Open Science Framework. AC acknowledges the support of the Agence
Nationale de la Recherche (ANR-17-CE28-0007 LangAge, ANR-16-DATA-0004,
ANR-14-CE30-0003, ANR-17-EURE-0017); and the J. S. McDonnell Foundation
Understanding Human Cognition Scholar Award.

Correspondence concerning this article should be addressed to
Alejandrina Cristia, 29, rue d'Ulm, 75005 Paris, France. E-mail:
\href{mailto:alecristia@gmail.com}{\nolinkurl{alecristia@gmail.com}}}

\abstract{
In nonword repetition (NWR) studies, participants are presented
auditorily with an item that is phonologically legal but lexically
meaningless in the local language. Accuracy is thought to reflect
long-term phonological knowledge as well as online phonological working
memory and flexible production patterns.In this study, we report on NWR
results among children learning Yêly Dnyé, an isolate spoken in Rossel
Island, PNG, with an unusually dense phonological inventory. This study
contributes to three lines of research. First, we document that non-word
items containing typologically rare sounds are repeated accurately less
often that non-words containing more common sounds. Second, we document
rather weak effects of item length, contributing to other research
suggesting that length effects may be language-specific. Third, we do
not find strong individual variation effects in this population,
contrary to previous results documenting strong age-related effects.
Together, these data provide a unique view of online phonological
processing in a seldom-studied language, and contribute to both
typological and language acquisition research.

}

@@

\subsection{TODO Middy}\label{todo-middy}

\begin{itemize}
\tightlist
\item
  look through this manuscript for bits of text to be written by
  \textbf{mc}
\item
  help with table 1 (formatting \& making it appear, because it's broken
  by the phonological representations)
\item
  do you think Figure 2 should be done with phonetic symbols? If yes,
  could you try to implement that?
\item
  double check demo -- probably start with the xlsx version that has the
  children's first names, date of birth, and date of test. I looked
  again at my local files (still haven't checked the labs' ones) and I'm
  surprised I have maternal education for more children than I have
  birth order for\ldots{} If you correct this, to generate the csv I
  also round the age, assign age.rounded to reported age (in years) when
  DOB is missing, remove first names, DOB, DOTest and age.precise before
  pushing to github
\item
  double check my frequency entries in segments.xlsx using the search
  function in \url{https://phoible.org/parameters}; take a look at
  \url{http://phoible.github.io/conventions/} in case you see something
  about double articulation (I didn't find tp or lBj)
\item
  double check phonological implementation of stimuli in stimuli.xlsx
\item
  for both of the above, if you make any changes, they need to be
  re-saved as UTF16 txt file, and then converted (see remember above)
\item
  double check also stim\_seg\_freq -- why are some items still with NA
  freq?
\item
  consent, ``correct'', ``accuracy'', ``performance''
\item
  draft discussion
\end{itemize}

\subsection{TODO later}\label{todo-later}

@@

\subsection{Introduction}\label{introduction}

@@

\subsubsection{Yélî Dnye phonological
overview}\label{yuxe9luxee-dnye-phonological-overview}

Yélî Dnye is an isolate language (presumed Papuan) spoken by
approximately 7000 people residing on Rossel Island, which is found at
the far end of the Louisiade Archipelago in Milne Bay Province, Papua
New Guinea. The Yélî sound system, much like its baroque grammatical
system (Levinson, in preparation), is unlike any other in the region.

With only four primary places of articulation (bilabial, alveolar,
post-alveolar, and velar) and no voicing contrasts, the phonological
inventory is remarkably packed with acoustically similar segments. The
core oral stop set includes both singleton (/p/, /t/, /ṭ/, and /k/) and
doubly-articulated (/tp/, /ṭp/, /kp/) segments, with full nasal
equivalents (/m/, /n/, /ṇ/, /ŋ/, /nm/, /ṇm/, /ŋm/), and with a
substantial portion of these able to be contrastively pre-nasalized or
nasally released (/mp/, /nt/, /ṇṭ/, /ŋk/, /nmtp/, /ṇmṭp/, /ŋmkp/, /ṭṇ/,
/kŋ/, /ṭpṇm/, /kpŋm/). A large number of this combinatorial set can
further be contrastively labialized or palatalized on release, or both
(e.g., /pʲ/, /pʷ/, /pʲʷ/; /tpʲ/; /ṇmḍbʲ/; see Levinson (in preparation)
for details). The consonantal inventory also includes a number of
non-nasal continuants (/w/, /j/, /ɣ/, /l/, /βʲ/, /lʲ/, /lβʲ/). Vowels in
Yélî Dnye may be oral or nasal, short or long. The 10 oral vowel
qualities, which span four levels of vowel height, (/i/, /ɯ/, /u/, /e/,
/o/, /ə/, /ɛ/, /ɔ/, /æ/, /ɑ/) can be produced as short or long vowels,
with 7 of these able to appear as short and long nasal vowels as well
(/i/, /u/, /ə/, /ɛ/, /ɔ/, /æ/, /ɑ/).

In total, Yélî Dnye uses 90 distinctive segments (93 when including
consonants that are extremely rarely used), far outstripping the
phonemic inventory size of other documented Papuan languages (Levinson
(in preparation); Foley 1986) and includes at least two contrasts not
yet been documented elsewhere {[}labial-coronal double-articulations
with dental vs.~post-alveolar coronal placement in both oral and nasal
stops, Maddieson \& Levinson 2001; Ladefoged and Maddieson 1996{]}.

Most words in Yélî Dnye are bisyllabic (\textasciitilde{}50\%), with
monosyllabic words (\textasciitilde{}40\%) appearing most commonly after
that, and with tri-and-above syllablic words appearing least frequently
(\textasciitilde{}10\%; based on \textgreater{}5800 lexemes in the
current Yélî Dnye dictionary, Levinson, in preparation). The vast
majority of syllables use a CV format. A small portion of the lexicon
features words with a final closed syllable, limited to codas of -/m/,
-/p/, or -/j/ (e.g., `ndap' (Spondylus shell) /ṇṭæp/). However, in
spontaneous speech an epenthetic /ɯ/ is often inserted after word-final
coda, yielding a CVCV structure instead (e.g., `ndapî' /`ṇṭæ.pɯ/). This
process is used frequently with English loan words that have a coda
(e.g., 'ponî' (phone) /`pɔ.nɯ/). An even smaller portion of the lexicon
features words starting with a vowel (e.g., 'ala' (here) /æ.'læ/), but
these are limited to /æ/-. Finally, the lexicon features a handful of
single-vowel grammatical morphemes (see Levinson (in preparation) for
details).

While all consonants can appear in the first syllable of the word, in
bisyllabic roots, the second syllable onset consonant is almost
exclusively drawn from a subset of the simplex consonants (in order of
descending frequency: /m/, /ṭ/, /n/, /p/, /l/, /w/, /j/, /pʷ/, /ŋ/, /k/,
/t/, /ɣ/, /mʷ/). In monosyllabic stems, oral vowels (short: 43\% and
long: 31\%) are more common than nasal ones (short: 4\% and long: 22\%),
with the first syllable of bisyllabic stems following a similar pattern.
Most vowels in the second syllable of bisyllabic stems are also short
and oral, with a potential preference for rounding harmony between the
first and second vowels. Stress typically falls on the first syllable of
bisyllabic words, on the first and third syllables of tetrasyllabic
words, and on the first or second syllable of trisyllabic words,
depending on its vocalic features (see Levinson (in preparation) for
details).

\subsubsection{\texorpdfstring{Intro to the people (\textbf{mc}) -
please feel free to throw away anything that is not
useful!}{Intro to the people (mc) - please feel free to throw away anything that is not useful!}}\label{intro-to-the-people-mc---please-feel-free-to-throw-away-anything-that-is-not-useful}

Most speakers of the language grow up speaking it monolingually within
their swidden horticulturalist communities until they begin attending
school around the age of 7; school instruction is in English. While
monolingual Yélî Dnye upbringing is common, multilingual families are
not unusual, particularly in the region around the Catholic Mission
(1953--present) where there appears to be a higher incidence of
married-in mothers from other islands (Brown \& Casillas, in press), the
same region in which the current data were collected. Children in these
multilingual families grow up speaking Yélî Dnye and English, Tok Pisin,
and/or some other language(s) from the region.

@@

\subsubsection{Brief review of NWR for our
purposes}\label{brief-review-of-nwr-for-our-purposes}

@@

\subsubsection{Research questions}\label{research-questions}

@@

\subsection{Methods}\label{methods}

\subsubsection{Stimuli}\label{stimuli}

@@

A first list of candidate items was generated in 2018 by selecting
simple consonants (``p'', ``t'', ``d'', ``k'', ``m'', ``n'', ``w'',
``y'') and vowels (``i'', ``o'',``u'', ``a'',``e'') that were combined
into consonant-vowel syllables, further sampling the space of 1- to
4-syllable sequences. These candidates were automatically checked
against Levinson's 2015 dictionary and removed from consideration if
they appeared in the dictionary. The second author presented them orally
to three local research assistants, who were asked to repeat them and
further say whether they were real words. Any item for which two or more
of the assistants reported them having a meaning or some form of
association was excluded.

A second list of candidate items was generated in 2019 by selecting
complex consonants and systematically crossing them with all the vowels
in the Yélî inventory to produce consonant-vowel monosyllables. As
before, items were automatically excluded if they appeared in the
dictionary. Additionally, since hearing vowel length in monosyllables in
isolation is challenging, any item that had a short/long real word
neighbor was filtered out. Since the phonology and phonetics of Yélî is
still in the process of being described {[}CITE \textbf{mc} please fill
in{]}, there could have been undocumented constraints that rendered
items illegal. Therefore, we made sure that the precise consonant-vowel
sequence occurred in some real word in the dictionary (i.e., that there
was a longer word included the monosyllable as a subsequence). These
candidates were presented to one informant, for a final check that they
did not mean anything. Together with the 2018 selection, they were
recorded using a headset \textbf{XXX }mc** please fill in** and an
Olympus \textbf{XXX }mc** please fill in** from the written form
presented together with the same item orally (by the second author). The
complete recorded list was finally presented to two more informants, who
could repeat all the items and who confirmed there were no real words.
Even so, there was one monosyllable that was often identified as a real
word (intended ``yî'' /yXX/; identified as ``yi'' /yi/, \emph{tree}).
This item is removed from analyses below.

@@

\subsubsection{Procedure}\label{procedure}

@@

We visited four different hamlets once, and attempted to test all
eligible children present at the time, to prevent the items `spreading'
through hearsay. In the first hamlet, we tested children in five
different places, with some children being tested inside a house and
others tested on the veranda. The complete list of places and the ways
in which they met the desiderata mentioned above can be found in the raw
data, available from online supplementary materials.

@@

\subsubsection{Coding}\label{coding}

@@

\subsubsection{Analyses}\label{analyses}

@@

\subsubsection{Participants}\label{participants}

@@

\subsection{Results}\label{results}

\subsubsection{Preliminary analyses}\label{preliminary-analyses}

@@

The overall NWR score was M = 61\% (SD = 15\%). Scores based on phonemes
are even higher M = 77\% (SD = 9\%). The phoneme-based normalized
Levenshtein distance was M = 23\% (SD = 9\%), meaning that about a fifth
of phonemes were were substituted, inserted, or deleted. Notice that the
normalized Levenshtein distance is the complement of phoneme-based
scores.

\subsubsection{NWR as a function of cross-linguistic phone
frequency}\label{nwr-as-a-function-of-cross-linguistic-phone-frequency}

@@

We could include 885 observations (\textbf{why -- shouldn't it be
1k??}), from 41 children producing in any given trial one of 41
potential target words. The analysis revealed a significant estimate for
the average cross-linguistic frequency of phones in the target words (ß
= 0.03, SE ß = 0.01, p = 0): Target words with phones found more
frequently across languages had a higher proportion of words that were
correctly repeated, as clear in Figure 2.

@@

@@

\subsubsection{Accuracy a function of item
length}\label{accuracy-a-function-of-item-length}

@@

\begin{verbatim}
## 
## 
## \begin{table}[tbp]
## \begin{center}
## \begin{threeparttable}
## \caption{\label{tab:tablength}NWR scores measured in whole word accuracy, phoneme-based accuracy, and normalized Levenshtein Distance, separately for the four stimuli lengths.}
## \begin{tabular}{lll}
## \toprule
## Word & \multicolumn{1}{c}{Phoneme} & \multicolumn{1}{c}{NLD}\\
## \midrule
## 47 (22) & 59 (17) & 41 (17)\\
## 79 (22) & 94 (7) & 6 (7)\\
## 78 (19) & 94 (6) & 6 (6)\\
## 75 (32) & 93 (11) & 7 (11)\\
## \bottomrule
## \end{tabular}
## \end{threeparttable}
## \end{center}
## \end{table}
\end{verbatim}

\subsubsection{Factor structuring individual
variation}\label{factor-structuring-individual-variation}

@@

\subsection{Discussion}\label{discussion}

@@

\subsection{Acknowledgments}\label{acknowledgments}

@@

\section{References}\label{references}

>>>>

\PassOptionsToPackage{unicode=true}{hyperref} % options for packages loaded elsewhere
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[english,,man,floatsintext]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provides euro and other symbols
\else % if luatex or xelatex
  \usepackage{unicode-math}
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\usepackage{hyperref}
\hypersetup{
            pdftitle={Non-word repetition in Yélî Dnye},
            pdfkeywords={phonology, non-word repetition, development},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}

@@

\setmainfont{Doulos SIL} % Set main font to Doulos SIL
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

% Table formatting
\usepackage{longtable}
\usepackage{lscape}
% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
\usepackage{multirow}		% Table styling
\usepackage{tabularx}		% Control Column width
\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
\usepackage{threeparttablex}            % Lets threeparttable work with longtable

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\begin{center}\begin{threeparttable}}
%   {\end{threeparttable}\end{center}\end{landscape}}
\newenvironment{lltable}{\begin{landscape}\begin{center}\begin{ThreePartTable}}{\end{ThreePartTable}\end{center}\end{landscape}}

% Enables adjusting longtable caption width to table width
% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

% \setlength{\parindent}{0.5in}
% \setlength{\parskip}{0pt plus 0pt minus 0pt}

% \usepackage{etoolbox}
\makeatletter
\patchcmd{\HyOrg@maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\makeatother
\shorttitle{NWR in Yélî Dnye}
\author{Alejandrina Cristia\textsuperscript{1}\ \& Marisa Casillas\textsuperscript{2}}
\affiliation{
\vspace{0.5cm}
\textsuperscript{1} Laboratoire de Sciences Cognitives et de Psycholinguistique, Département d'Etudes cognitives, ENS, EHESS, CNRS, PSL University\\\textsuperscript{2} Max Planck Institute for Psycholinguistics}
\authornote{All data are made available in a repository in the Open Science Framework. AC acknowledges the support of the Agence Nationale de la Recherche (ANR-17-CE28-0007 LangAge, ANR-16-DATA-0004, ANR-14-CE30-0003, ANR-17-EURE-0017); and the J. S. McDonnell Foundation Understanding Human Cognition Scholar Award.

Correspondence concerning this article should be addressed to Alejandrina Cristia, 29, rue d’Ulm, 75005 Paris, France. E-mail: alecristia@gmail.com}
\keywords{phonology, non-word repetition, development\newline\indent Word count: ~xxx words}
\usepackage{lineno}

\linenumbers
\usepackage{csquotes}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[shorthands=off,main=english]{babel}
\else
  % load polyglossia as late as possible as it *could* call bidi if RTL lang (e.g. Hebrew or Arabic)
  \usepackage{polyglossia}
  \setmainlanguage[]{english}
\fi

\title{Non-word repetition in Yélî Dnye}

\date{}

\abstract{
In nonword repetition (NWR) studies, participants are presented auditorily with an item that is phonologically legal but lexically meaningless in the local language. Accuracy is thought to reflect long-term phonological knowledge as well as online phonological working memory and flexible production patterns.In this study, we report on NWR results among children learning Yêly Dnyé, an isolate spoken in Rossel Island, PNG, with an unusually dense phonological inventory. This study contributes to three lines of research. First, we document that non-word items containing typologically rare sounds are repeated accurately less often that non-words containing more common sounds. Second, we document rather weak effects of item length, contributing to other research suggesting that length effects may be language-specific. Third, we do not find strong individual variation effects in this population, contrary to previous results documenting strong age-related effects. Together, these data provide a unique view of online phonological processing in a seldom-studied language, and contribute to both typological and language acquisition research.
}

@@

\hypertarget{todo-middy}{%
\subsection{TODO Middy}\label{todo-middy}}

\begin{itemize}
\tightlist
\item
  look through this manuscript for bits of text to be written by \textbf{mc}
\item
  help with table 1 (formatting \& making it appear, because it's broken by the phonological representations)
\item
  do you think Figure 2 should be done with phonetic symbols? If yes, could you try to implement that?
\item
  double check demo -- probably start with the xlsx version that has the children's first names, date of birth, and date of test. I looked again at my local files (still haven't checked the labs' ones) and I'm surprised I have maternal education for more children than I have birth order for\ldots{} If you correct this, to generate the csv I also round the age, assign age.rounded to reported age (in years) when DOB is missing, remove first names, DOB, DOTest and age.precise before pushing to github
\item
  double check my frequency entries in segments.xlsx using the search function in \url{https://phoible.org/parameters}; take a look at \url{http://phoible.github.io/conventions/} in case you see something about double articulation (I didn't find tp or lBj)
\item
  double check phonological implementation of stimuli in stimuli.xlsx
\item
  for both of the above, if you make any changes, they need to be re-saved as UTF16 txt file, and then converted (see remember above)
\item
  double check also stim\_seg\_freq -- why are some items still with NA freq?
\item
  consent, \enquote{correct}, \enquote{accuracy}, \enquote{performance}
\item
  draft discussion
\end{itemize}

\hypertarget{todo-later}{%
\subsection{TODO later}\label{todo-later}}

@@

\hypertarget{introduction}{%
\subsection{Introduction}\label{introduction}}

@@

\hypertarget{yuxe9luxee-dnye-phonological-overview}{%
\subsubsection{Yélî Dnye phonological overview}\label{yuxe9luxee-dnye-phonological-overview}}

Yélî Dnye is an isolate language (presumed Papuan) spoken by approximately 7000 people residing on Rossel Island, which is found at the far end of the Louisiade Archipelago in Milne Bay Province, Papua New Guinea. The Yélî sound system, much like its baroque grammatical system (Levinson, n.d.), is unlike any other in the region.

With only four primary places of articulation (bilabial, alveolar, post-alveolar, and velar) and no voicing contrasts, the phonological inventory is remarkably packed with acoustically similar segments. The core oral stop set includes both singleton (/p/, /t/, /ṭ/, and /k/) and doubly-articulated (/tp/, /ṭp/, /kp/) segments, with full nasal equivalents (/m/, /n/, /ṇ/, /ŋ/, /nm/, /ṇm/, /ŋm/), and with a substantial portion of these able to be contrastively pre-nasalized or nasally released (/mp/, /nt/, /ṇṭ/, /ŋk/, /nmtp/, /ṇmṭp/, /ŋmkp/, /ṭṇ/, /kŋ/, /ṭpṇm/, /kpŋm/). A large number of this combinatorial set can further be contrastively labialized or palatalized on release, or both (e.g., /pʲ/, /pʷ/, /pʲʷ/; /tpʲ/; /ṇmḍbʲ/; see Levinson (n.d.) for details). The consonantal inventory also includes a number of non-nasal continuants (/w/, /j/, /ɣ/, /l/, /βʲ/, /lʲ/, /lβʲ/). Vowels in Yélî Dnye may be oral or nasal, short or long. The 10 oral vowel qualities, which span four levels of vowel height, (/i/, /ɯ/, /u/, /e/, /o/, /ə/, /ɛ/, /ɔ/, /æ/, /ɑ/) can be produced as short or long vowels, with 7 of these able to appear as short and long nasal vowels as well (/i/, /u/, /ə/, /ɛ/, /ɔ/, /æ/, /ɑ/).

In total, Yélî Dnye uses 90 distinctive segments (93 when including consonants that are extremely rarely used), far outstripping the phonemic inventory size of other documented Papuan languages (Levinson (n.d.); Foley 1986) and includes at least two contrasts not yet been documented elsewhere {[}labial-coronal double-articulations with dental vs.~post-alveolar coronal placement in both oral and nasal stops, Maddieson \& Levinson 2001; Ladefoged and Maddieson 1996{]}.

Most words in Yélî Dnye are bisyllabic (\textasciitilde{}50\%), with monosyllabic words (\textasciitilde{}40\%) appearing most commonly after that, and with tri-and-above syllablic words appearing least frequently (\textasciitilde{}10\%; based on \textgreater{}5800 lexemes in the current Yélî Dnye dictionary, Levinson, n.d.). The vast majority of syllables use a CV format. A small portion of the lexicon features words with a final closed syllable, limited to codas of -/m/, -/p/, or -/j/ (e.g., \enquote{ndap} (Spondylus shell) /ṇṭæp/). However, in spontaneous speech an epenthetic /ɯ/ is often inserted after word-final coda, yielding a CVCV structure instead (e.g., \enquote{ndapî} /\enquote{ṇṭæ.pɯ/). This process is used frequently with English loan words that have a coda (e.g., 'ponî} (phone) /\enquote{pɔ.nɯ/). An even smaller portion of the lexicon features words starting with a vowel (e.g., 'ala} (here) /æ.'læ/), but these are limited to /æ/-. Finally, the lexicon features a handful of single-vowel grammatical morphemes (see Levinson (n.d.) for details).

While all consonants can appear in the first syllable of the word, in bisyllabic roots, the second syllable onset consonant is almost exclusively drawn from a subset of the simplex consonants (in order of descending frequency: /m/, /ṭ/, /n/, /p/, /l/, /w/, /j/, /pʷ/, /ŋ/, /k/, /t/, /ɣ/, /mʷ/). In monosyllabic stems, oral vowels (short: 43\% and long: 31\%) are more common than nasal ones (short: 4\% and long: 22\%), with the first syllable of bisyllabic stems following a similar pattern. Most vowels in the second syllable of bisyllabic stems are also short and oral, with a potential preference for rounding harmony between the first and second vowels. Stress typically falls on the first syllable of bisyllabic words, on the first and third syllables of tetrasyllabic words, and on the first or second syllable of trisyllabic words, depending on its vocalic features (see Levinson (n.d.) for details).

\hypertarget{intro-to-the-people-mc---please-feel-free-to-throw-away-anything-that-is-not-useful}{%
\subsubsection{\texorpdfstring{Intro to the people (\textbf{mc}) - please feel free to throw away anything that is not useful!}{Intro to the people (mc) - please feel free to throw away anything that is not useful!}}\label{intro-to-the-people-mc---please-feel-free-to-throw-away-anything-that-is-not-useful}}

Most speakers of the language grow up speaking it monolingually within their swidden horticulturalist communities until they begin attending school around the age of 7; school instruction is in English. While monolingual Yélî Dnye upbringing is common, multilingual families are not unusual, particularly in the region around the Catholic Mission (1953--present) where there appears to be a higher incidence of married-in mothers from other islands (Brown \& Casillas, n.d.), the same region in which the current data were collected. Children in these multilingual families grow up speaking Yélî Dnye and English, Tok Pisin, and/or some other language(s) from the region.

@@

\hypertarget{brief-review-of-nwr-for-our-purposes}{%
\subsubsection{Brief review of NWR for our purposes}\label{brief-review-of-nwr-for-our-purposes}}

@@

\hypertarget{research-questions}{%
\subsubsection{Research questions}\label{research-questions}}

@@

\hypertarget{methods}{%
\subsection{Methods}\label{methods}}

\hypertarget{stimuli}{%
\subsubsection{Stimuli}\label{stimuli}}

@@

A first list of candidate items was generated in 2018 by selecting simple consonants (\enquote{p}, \enquote{t}, \enquote{d}, \enquote{k}, \enquote{m}, \enquote{n}, \enquote{w}, \enquote{y}) and vowels (\enquote{i}, \enquote{o},\enquote{u}, \enquote{a},\enquote{e}) that were combined into consonant-vowel syllables, further sampling the space of 1- to 4-syllable sequences. These candidates were automatically checked against Levinson's 2015 dictionary and removed from consideration if they appeared in the dictionary. The second author presented them orally to three local research assistants, who were asked to repeat them and further say whether they were real words. Any item for which two or more of the assistants reported them having a meaning or some form of association was excluded.

A second list of candidate items was generated in 2019 by selecting complex consonants and systematically crossing them with all the vowels in the Yélî inventory to produce consonant-vowel monosyllables. As before, items were automatically excluded if they appeared in the dictionary. Additionally, since hearing vowel length in monosyllables in isolation is challenging, any item that had a short/long real word neighbor was filtered out. Since the phonology and phonetics of Yélî is still in the process of being described {[}CITE \textbf{mc} please fill in{]}, there could have been undocumented constraints that rendered items illegal. Therefore, we made sure that the precise consonant-vowel sequence occurred in some real word in the dictionary (i.e., that there was a longer word included the monosyllable as a subsequence). These candidates were presented to one informant, for a final check that they did not mean anything. Together with the 2018 selection, they were recorded using a headset \textbf{XXX }mc** please fill in** and an Olympus \textbf{XXX }mc** please fill in** from the written form presented together with the same item orally (by the second author). The complete recorded list was finally presented to two more informants, who could repeat all the items and who confirmed there were no real words. Even so, there was one monosyllable that was often identified as a real word (intended \enquote{yî} /yXX/; identified as \enquote{yi} /yi/, \emph{tree}). This item is removed from analyses below.

@@

\hypertarget{procedure}{%
\subsubsection{Procedure}\label{procedure}}

@@

We visited four different hamlets once, and attempted to test all eligible children present at the time, to prevent the items \enquote{spreading} through hearsay. In the first hamlet, we tested children in five different places, with some children being tested inside a house and others tested on the veranda. The complete list of places and the ways in which they met the desiderata mentioned above can be found in the raw data, available from online supplementary materials.

@@

\hypertarget{coding}{%
\subsubsection{Coding}\label{coding}}

@@

\hypertarget{analyses}{%
\subsubsection{Analyses}\label{analyses}}

@@

\hypertarget{participants}{%
\subsubsection{Participants}\label{participants}}

@@

\hypertarget{results}{%
\subsection{Results}\label{results}}

\hypertarget{preliminary-analyses}{%
\subsubsection{Preliminary analyses}\label{preliminary-analyses}}

@@

The overall NWR score was M = 61\% (SD = 15\%). Scores based on phonemes are even higher M = 78\% (SD = 9\%). The phoneme-based normalized Levenshtein distance was M = 22\% (SD = 9\%), meaning that about a fifth of phonemes were were substituted, inserted, or deleted. Notice that the normalized Levenshtein distance is the complement of phoneme-based scores.

\hypertarget{nwr-as-a-function-of-cross-linguistic-phone-frequency}{%
\subsubsection{NWR as a function of cross-linguistic phone frequency}\label{nwr-as-a-function-of-cross-linguistic-phone-frequency}}

@@

We could include 885 observations (\textbf{why -- shouldn't it be 1k??}), from 41 children producing in any given trial one of 41 potential target words. The analysis revealed a significant estimate for the average cross-linguistic frequency of phones in the target words (ß = 0.77, SE ß = 0.19, p = 0): Target words with phones found more frequently across languages had a higher proportion of words that were correctly repeated, as clear in Figure 2. Age was also a significant predictor (ß = 0.31, SE ß = 0.13, p = 0.02).

@@

We next checked whether this may actually be due to frequency of the sounds within the language. We estimated frequency of the phones present in the stimuli in a corpus of child-centered recordings ({\textbf{???}}), and applied the natural logarithm. As before, unattested sounds were not considered (i.e., they were declared NA so that they do not count for analyses). Phone frequency estimated from this corpus were correlated with cross-linguistic phone frequencies (r(23)=0.65, =0). Therefore, we fit another mixed logistic regression, this time declaring as fixed effects scaled cross-linguistic and corpus frequencies (averaged across all attested phones within each stimulus item), in addition to age. As before, the model contained random slopes for both child ID and target. In this model, both cross-linguistic frequency (ß = 0.50, SE ß = 0.25, p = 0.04) and age (ß = 0.31, SE ß = 0.13, p = 0.02) were significant predictors of accuracy, but corpus frequency (ß = 0.37, SE ß = 0.24, p = 0.12) was not.

@@

\hypertarget{accuracy-as-a-function-of-item-length}{%
\subsubsection{Accuracy as a function of item length}\label{accuracy-as-a-function-of-item-length}}

@@

\begin{table}[tbp]

\begin{center}
\begin{threeparttable}

\caption{\label{tab:tablength}NWR scores measured in whole word accuracy, phoneme-based accuracy, and normalized Levenshtein Distance, separately for the four stimuli lengths.}

\begin{tabular}{lll}
\toprule
Word & \multicolumn{1}{c}{Phoneme} & \multicolumn{1}{c}{NLD}\\
\midrule
47 (22) & 59 (17) & 41 (17)\\
79 (22) & 94 (7) & 6 (7)\\
78 (19) & 94 (6) & 6 (6)\\
75 (32) & 93 (11) & 7 (11)\\
\bottomrule
\end{tabular}

\end{threeparttable}
\end{center}

\end{table}

\hypertarget{factor-structuring-individual-variation}{%
\subsubsection{Factor structuring individual variation}\label{factor-structuring-individual-variation}}

@@

\hypertarget{discussion}{%
\subsection{Discussion}\label{discussion}}

@@

\hypertarget{acknowledgments}{%
\subsection{Acknowledgments}\label{acknowledgments}}

@@

\hypertarget{references}{%
\section{References}\label{references}}
